Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2018-01-30 paper.tex
Remove some out-of-date points

\begin{document}
\title{Motional Ground State Cooling Outside the Lamb-Dicke Regime}
\author{Y. Yu}
\email{yichaoyu@g.harvard.edu}
\author{N. R. Hutzler}
\altaffiliation{Present address: California Institute of Technology, Division of Physics, Mathematics, and Astronomy. Pasadena, CA, 91125}
\author{J. T. Zhang}
\author{L. R. Liu}
\author{T. Rosenband}
\author{K.-K. Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

@@

There are two additional important challenges in cooling single sodium atoms.
First, the initial temperature populates high motional states,
causing the atoms to sample the anharmonicity of the trap away from the center.
Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$
for each trap axis $i$, and calculated from the quartic term
of the optical tweezers via perturbation theory.
In the paraxial approximation, we find $A_{i,n}=\frac{-3n\hbar}{4\pi m d_i^2}$,
where $d_i$ equals the beam radius for the radial directions and
$d_z\approx\pi w_{0,x}w_{0,y}/\lambda_{\textrm{trap}}$.
Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{-1.4, -1.4, -0.16\}n$~kHz.
For mixed states, this broadens high-order sidebands due to the $n$-dependence of the transitions.
In order to mitigate this, we drive the sidebands with a large Rabi frequency
and short pulses to Fourier broaden the spectrum.
However, anharmonicity may still shift high-$n$ states out of resonance
and prevent effective cooling.

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $540$ cooling pulses (total duration $53$ ms) are applied
along three axes with cooling beginning on the radial second order and axial eighth order.
The full RSC pulse sequence is given in~\footnote{See supplemental material}.
To characterize the single atom thermal state before and after cooling,
we perform Raman sideband thermometry~\cite{Monroe1995, Meekhof1996}.
For the more tightly confined (non-degenerate) radial directions,
we observe clear $\Delta n=1$, $\Delta n=-1$, and $\Delta n=-2$ sidebands before RSC,
as is shown in Fig.~\ref{f-radial}A.
After RSC, the $\Delta n=-1$ and $\Delta n=-2$ sidebands on both radial axes are strongly reduced.
The formula that equates the ratio of $\Delta n=-1$ and $\Delta n=1$ sideband heights to
$\bar{n}/(\bar{n}+1)$ ($\bar{n}$ is the mean motional quantum number of the assumed thermal state)
was derived in the LD regime~\cite{Monroe1995}.
However, it is also valid outside the LD regime.
The resulting $\bar{n}_x=0.019(4)$ and $\bar{n}_y=0.024(3)$ correspond to
ground-state fractions of 98.1(5)\% and 97.6(3)\%,
in agreement with fitted values of 98(1)\% and 95(3)\%
from the Rabi flopping curves~\cite{Meekhof1996} in Fig.~\ref{f-radial}B and \ref{f-radial}C.
The initial temperature of $80\,\mu$K before RSC is obtained from similar fits.

@@

Combining the axial and radial cooling results,
a single Na atom is in the 3D ground state with a probability of 93.5(7)\% after RSC.
We note that there is an additional 4\% population loss from imaging.
The cooling efficiency is limited by spontaneous scattering rate
(0.1-0.2 kHz) from the Raman beams,
as well as spectral broadening from magnetic field fluctuations and trap anharmonicity.

@@

We expect to enhance the ground state probability after RSC
by increasing the detuning of the Raman beams,
addressing a wider range of axial trap anharmonicity,
implementing better control of the magnetic field, and optimizing imaging.
Another improvement could be to implement grey molasses cooling to achieve
a lower starting temperature before RSC~\cite{Colzi2016}.

@@

>>>>

\begin{document}
\title{Motional Ground State Cooling Outside the Lamb-Dicke Regime}
\author{Y. Yu}
\email{yichaoyu@g.harvard.edu}
\author{N. R. Hutzler}
\altaffiliation{Present address: California Institute of Technology, Division of Physics, Mathematics, and Astronomy. Pasadena, CA, 91125}
\author{J. T. Zhang}
\author{L. R. Liu}
\author{J. D. Hood}
\author{T. Rosenband}
\author{K.-K. Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

@@

There are two additional important challenges in cooling single sodium atoms.
First, the initial temperature populates high motional states,
causing the atoms to sample the anharmonicity of the trap away from the center.
Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$
for each trap axis $i$, and calculated from the quartic term
of the optical tweezers via perturbation theory.
In the paraxial approximation, we find $A_{i,n}=\frac{-3n\hbar}{4\pi m d_i^2}$,
where $d_i$ equals the beam radius for the radial directions and
$d_z\approx\pi w_{0,x}w_{0,y}/\lambda_{\textrm{trap}}$.
Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{-1.4, -1.4, -0.16\}n$~kHz.
For mixed states, this broadens high-order sidebands due to the $n$-dependence of the transitions.
In order to mitigate this, we drive the sidebands with a large Rabi frequency
and short pulses to Fourier broaden the spectrum.

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $540$ cooling pulses (total duration $53$ ms) are applied
along three axes with cooling beginning on the radial second order and axial fifth order.
The full RSC pulse sequence is given in~\footnote{See supplemental material}.
To characterize the single atom thermal state before and after cooling,
we perform Raman sideband thermometry~\cite{Monroe1995, Meekhof1996}.
For the more tightly confined radial directions,
we observe clear $\Delta n=1$, $\Delta n=-1$, and $\Delta n=-2$ sidebands before RSC,
as is shown in Fig.~\ref{f-radial}A.
After RSC, the $\Delta n=-1$ and $\Delta n=-2$ sidebands on both radial axes are strongly reduced.
The formula that equates the ratio of $\Delta n=-1$ and $\Delta n=1$ sideband heights to
$\bar{n}/(\bar{n}+1)$ ($\bar{n}$ is the mean motional quantum number of the assumed thermal state)
was derived in the LD regime~\cite{Monroe1995}.
However, it is also valid outside the LD regime.
The resulting $\bar{n}_x=0.019(4)$ and $\bar{n}_y=0.024(3)$ correspond to
ground-state fractions of 98.1(5)\% and 97.6(3)\%,
in agreement with fitted values of 98(1)\% and 95(3)\%
from the Rabi flopping curves~\cite{Meekhof1996} in Fig.~\ref{f-radial}B and \ref{f-radial}C.
The initial temperature of $80\,\mu$K before RSC is obtained from similar fits.

@@

Combining the axial and radial cooling results,
a single Na atom is in the 3D ground state with a probability of 93.5(7)\% after RSC.
We note that there is an additional 4\% population loss from imaging.
The cooling efficiency is limited by spontaneous scattering rate
(0.1-0.2 kHz) from the Raman beams,
as well as spectral broadening from magnetic field fluctuations.

@@

We expect to enhance the ground state probability after RSC
by increasing the detuning of the Raman beams,
implementing better control of the magnetic field, and optimizing imaging.
Another improvement could be to implement grey molasses cooling to achieve
a lower starting temperature before RSC~\cite{Colzi2016}.

@@

%% TODO
% Plot marker
% Remove plot title
% Remove mention of imaging loss in the text and move it to the figure caption
% Add 96% survival line.
% 94(1) ?
